## rOpenSci Demo

[rOpenSci](http://ropensci.org/)

### Getting data from the literature - PLOS

This example demonstrates how you can easily get literature data from Public Library of Science from R. 

#### Install rplos

```{r install, eval=FALSE}
install.packages(c("rplos","tm"))
```

#### Load rplos

```{r load}
library(rplos)
```

#### Search for mentions of Fisher in the author field, returning title and author fields, searching in full papers (not including figure captions, etc.), returning only 25 results.

```{r searchplos, message=FALSE, warning=FALSE}
key <- "key"
out <- searchplos(terms='author:"Fisher"', fields='title,author', toquery='doc_type:full', limit=25, key=key)
head(out) # first six rows
```

#### Quickly visualize variation in frequency of word usage in PLOS journals

```{r plosword, message=FALSE, warning=FALSE}
library(ggplot2)
plosword(list('monkey','Helianthus','sunflower','protein','whale'), key=key, vis = TRUE)$plot + 
  theme_grey(base_size=18)
```

#### Get abstracts of 500 papers, and use the tm package for text mining. 

Get 500 abstracts from PLOS One only. The `*:*` is special syntax to denote *give back everything*

```{r getabstracts, message=FALSE, warning=FALSE}
out <- searchplos(terms='*:*', fields='abstract', toquery=list('cross_published_journal_key:PLoSONE', 'doc_type:full'), limit=500, key=key)
out$abstract[1:3] # take a peek
```

Load the tm package, and create a document library

```{r corpus, message=FALSE, warning=FALSE}
library(tm)
(corpus <- Corpus(VectorSource(out$abstract)))
```

Create a term-document matrix from the corpus, and inspect it. 

```{r tdm, message=FALSE, warning=FALSE}
tdm <- DocumentTermMatrix(corpus,
                          control = list(removePunctuation = TRUE,
                                         stopwords = TRUE, 
                                         removeNumbers = TRUE))
inspect(tdm[1:5,1:5])
```

Various operations on document term matrices

```{r freqterm, message=FALSE, warning=FALSE}
# find terms that occur at least five times across documents
findFreqTerms(tdm, 250)

# find associations (terms which correlate) with at least 0.3 correlation for the term "result"
findAssocs(tdm, "cells", 0.3)
```